HO unification from object language to meta language

### 2024-02-16 main.tex
wip abstract

\newcommand*{\eqtau}{\ensuremath{\mathrel{\overset{\mathrm{\tau}}{=}}}}
\begin{abstract}

@@

\end{abstract}

@@

>>>>

\newcommand*{\eqtau}{\ensuremath{\mathrel{\overset{\mathrm{\tau}}{=}}}}

\begin{comment}
\begin{abstract}

@@

\end{abstract}
\end{comment}

\begin{abstract}
Specifying and implementing a logic from scratch requires a
considerable effort. Logical frameworks and Higher Order Logic Programming
Languages provide a reusable
Higher Order Meta Languages (ML) to make that easier:
the specification and implementation of binding and substitution comes for free 
if one reuses the the ML binders for the Object Logic (OL);
similarly proof construction (and even proof search) is greatly simplified
by leveraging the unification of the ML.
Notable examples of ML are Elf~\cite{elf}, Twelf~\cite{twelf},
$\lambda$Prolog~\cite{miller_nadathur_2012} and Isabelle~\cite{10.1007/978-3-540-71067-7_7}
that were used to implement various OLs such as
First Order Logic~\cite{felty88cade},
Set Theory~\cite{10.1007/BF00881873} and
even Higher Order Logic~\cite{books/sp/NipkowPW02}.

We are interested in using a ML to describe Dependent Type Theory (TT) and
implement a proof search procedure based on the unification (U) of the ML.
U is typically able to solve higher order problems in the pattern
fragment~\cite{miller92jsc} and identify ML terms up to $\eta\beta$.
The problem we face is that a natural encoding of the OL (FOOL) force on
U a ``first order''.
To overcome the problem we provide an alternative encoding of TT (HOOL) where
higher order unification problems can be faithfully delegated to U.
We show how to map FOOL to HOOL and how to map back the
unifiers found by U, effectively implementing an higher order unification
algorithm for FOOL.

We apply this technique to the implementation of a type-class solver for Coq,
which is a proof search procedure based on back-chaining designated lemmas.
Type-class solvers provide essential automation to widely used 
Coq libraries such as Stdpp/Iris~\cite{JUNG_KREBBERS_JOURDAN_BIZJAK_BIRKEDAL_DREYER_2018}
and TLC~\cite{10.1007/978-3-642-14052-5_15} that constitute our test bed.
\end{abstract}

@@

related~\cite{felty93lics}
